---
title: "Attrition Project"
author: "Steven Mann"
date: "8/13/2020"
output:
  html_document: default
  word_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

##Executive Summary
## This is a presentation on predicting employee turnover in Frito Lay. An interaction between
## Job Role & OverTime was determined to be the best indicator in predicting employee turnover rates.


#Read in the Data
```{r Read in Attrition Files}
CaseStudy = read.csv(file.choose(),header=TRUE,stringsAsFactors = TRUE)
NoAttrition = read.csv(file.choose(),header = TRUE, stringsAsFactors = TRUE)
NoSalary = read.csv(file.choose(),header = TRUE, stringsAsFactors = TRUE)
```

#Load in the Libraries
```{r Load Libraries}
library(tidyverse)
library(ggplot2)
library(caret)
library(e1071)
library(leaps)
library(stats)
library(rsq)
library(Metrics)
library(kknn)
```


##Create factor variables out of variables brought in as integers
```{r create factor variables}
names <- c('RelationshipSatisfaction','PerformanceRating','WorkLifeBalance',
           'JobInvolvement','JobSatisfaction','JobLevel','StockOptionLevel',
           'Education','EnvironmentSatisfaction')

CaseStudy[,names] <- lapply(CaseStudy[,names],factor,ordered = TRUE)
NoAttrition[,names] <- lapply(NoAttrition[,names],factor,ordered = TRUE)
NoSalary[,names] <- lapply(NoSalary[,names],factor,ordered = TRUE)
```

#Create new variable for job OT
```{r Job OT}
CaseStudy <- CaseStudy %>% mutate(JobRole_OT = paste(JobRole,OverTime,sep = "_"))
CaseStudy$JobRole_OT = as.factor(CaseStudy$JobRole_OT)
NoSalary <- NoSalary %>% mutate(JobRole_OT = paste(JobRole,OverTime,sep = "_"))
NoSalary$JobRole_OT = as.factor(NoSalary$JobRole_OT)
```

#Create categorical variable for Monthly Income based on summary statistics
```{r Monthly Income High/Low}
IncomeSum = summary(CaseStudy$MonthlyIncome)
CaseStudy$MonthlyIncomeHL = as.factor(if_else(CaseStudy$MonthlyIncome <= IncomeSum[2],"Lowest",
                                    if_else(CaseStudy$MonthlyIncome <= IncomeSum[3],"Medium/Low",
                                            if_else(CaseStudy$MonthlyIncome <= IncomeSum[4],"Medium/High","High"))))
```

#Create a categorical variable for Monthly Income based on summary statistics
```{r Totla Work High/Low}
TotalWorkSum = summary(CaseStudy$TotalWorkingYears)
CaseStudy$TotalWorkHL = as.factor(if_else(CaseStudy$TotalWorkingYears <= TotalWorkSum[2],"<=6",
                                  if_else(CaseStudy$TotalWorkingYears <= TotalWorkSum[3],"6< <=10",
                                          if_else(CaseStudy$TotalWorkingYears <= TotalWorkSum[5],"10< <=15",">15"))))
```

```{r create case study clean for models}
CaseStudyClean <- CaseStudy[,c("Attrition","JobRole_OT","MonthlyIncomeHL","JobInvolvement","WorkLifeBalance","TotalWorkHL",
                            "JobLevel","StockOptionLevel","MaritalStatus","Gender")]
```

##Pie chart for overall attrition rate
```{r JobRole_OT Baseline Chart}
data <- data.frame(a=CaseStudy$Attrition)
data <- data %>% 
  group_by(a) %>% 
  count() %>% 
  ungroup() %>% 
  mutate(per=`n`/sum(`n`)) %>% 
  arrange(desc(a))
data$label <- scales::percent(data$per)
ggplot(data=data)+
  geom_bar(aes(x="", y=per, fill=a), stat="identity", width = 1)+
  coord_polar("y", start=0)+
  theme_void()+
  geom_text(aes(x=1, y = cumsum(per) - per/2, label=label)) +
  labs(title = "Attrition Rates: Overall",fill = "Attrition")
```

#Split into Job Involvement Subsets for Plotting
```{r Job Involvement}
Involvement1 = CaseStudy %>% filter(JobInvolvement == 1)
Involvement2 = CaseStudy %>% filter(JobInvolvement == 2)
Involvement3 = CaseStudy %>% filter(JobInvolvement == 3)
Involvement4 = CaseStudy %>% filter(JobInvolvement == 4)

#Plot Pie chart for Job Involvement = 1
data <- data.frame(a=Involvement1$Attrition)
data <- data %>% 
  group_by(a) %>% 
  count() %>% 
  ungroup() %>% 
  mutate(per=`n`/sum(`n`)) %>% 
  arrange(desc(a))
data$label <- scales::percent(data$per)
ggplot(data=data)+
  geom_bar(aes(x="", y=per, fill=a), stat="identity", width = 1)+
  coord_polar("y", start=0)+
  theme_void()+
  geom_text(aes(x=1, y = cumsum(per) - per/2, label=label)) +
  labs(title = "Attrition Rates: Job Involvement = 1",fill = "Attrition")

#Plot Pie chart for Job Involvement = 2
data <- data.frame(a=Involvement2$Attrition)
data <- data %>% 
  group_by(a) %>% 
  count() %>% 
  ungroup() %>% 
  mutate(per=`n`/sum(`n`)) %>% 
  arrange(desc(a))
data$label <- scales::percent(data$per)
ggplot(data=data)+
  geom_bar(aes(x="", y=per, fill=a), stat="identity", width = 1)+
  coord_polar("y", start=0)+
  theme_void()+
  geom_text(aes(x=1, y = cumsum(per) - per/2, label=label)) +
  labs(title = "Attrition Rates: Job Involvement = 2",fill = "Attrition")

#Plot Pie chart for Job Involvement = 3
data <- data.frame(a=Involvement3$Attrition)
data <- data %>% 
  group_by(a) %>% 
  count() %>% 
  ungroup() %>% 
  mutate(per=`n`/sum(`n`)) %>% 
  arrange(desc(a))
data$label <- scales::percent(data$per)
ggplot(data=data)+
  geom_bar(aes(x="", y=per, fill=a), stat="identity", width = 1)+
  coord_polar("y", start=0)+
  theme_void()+
  geom_text(aes(x=1, y = cumsum(per) - per/2, label=label)) +
  labs(title = "Attrition Rates: Job Involvement = 3",fill = "Attrition")

#Plot Pie chart for Job Involvement = 4
data <- data.frame(a=Involvement4$Attrition)
data <- data %>% 
  group_by(a) %>% 
  count() %>% 
  ungroup() %>% 
  mutate(per=`n`/sum(`n`)) %>% 
  arrange(desc(a))
data$label <- scales::percent(data$per)
ggplot(data=data)+
  geom_bar(aes(x="", y=per, fill=a), stat="identity", width = 1)+
  coord_polar("y", start=0)+
  theme_void()+
  geom_text(aes(x=1, y = cumsum(per) - per/2, label=label)) +
  labs(title = "Attrition Rates: Job Involvement = 4",fill = "Attrition")
```

#Split into Job Level Subsets for Plotting
```{r Job Level Data}
Level1 = CaseStudy %>% filter(JobLevel == 1)
Level2 = CaseStudy %>% filter(JobLevel == 2)
Level3 = CaseStudy %>% filter(JobLevel == 3)
Level4 = CaseStudy %>% filter(JobLevel == 4)
Level5 = CaseStudy %>% filter(JobLevel == 5)

#Plot Pie chart for Job Level = 1
data <- data.frame(a=Level1$Attrition)
data <- data %>% 
  group_by(a) %>% 
  count() %>% 
  ungroup() %>% 
  mutate(per=`n`/sum(`n`)) %>% 
  arrange(desc(a))
data$label <- scales::percent(data$per)
ggplot(data=data)+
  geom_bar(aes(x="", y=per, fill=a), stat="identity", width = 1)+
  coord_polar("y", start=0)+
  theme_void()+
  geom_text(aes(x=1, y = cumsum(per) - per/2, label=label)) +
  labs(title = "Attrition Rates: Job Level = 1",fill = "Attrition")

#Plot Pie chart for Job Level = 2
data <- data.frame(a=Level2$Attrition)
data <- data %>% 
  group_by(a) %>% 
  count() %>% 
  ungroup() %>% 
  mutate(per=`n`/sum(`n`)) %>% 
  arrange(desc(a))
data$label <- scales::percent(data$per)
ggplot(data=data)+
  geom_bar(aes(x="", y=per, fill=a), stat="identity", width = 1)+
  coord_polar("y", start=0)+
  theme_void()+
  geom_text(aes(x=1, y = cumsum(per) - per/2, label=label)) +
  labs(title = "Attrition Rates: Job Level = 2",fill = "Attrition")

#Plot Pie chart for Job Level = 3
data <- data.frame(a=Level3$Attrition)
data <- data %>% 
  group_by(a) %>% 
  count() %>% 
  ungroup() %>% 
  mutate(per=`n`/sum(`n`)) %>% 
  arrange(desc(a))
data$label <- scales::percent(data$per)
ggplot(data=data)+
  geom_bar(aes(x="", y=per, fill=a), stat="identity", width = 1)+
  coord_polar("y", start=0)+
  theme_void()+
  geom_text(aes(x=1, y = cumsum(per) - per/2, label=label)) +
  labs(title = "Attrition Rates: Job Level = 3",fill = "Attrition")

#Plot Pie chart for Job Level = 4
data <- data.frame(a=Level4$Attrition)
data <- data %>% 
  group_by(a) %>% 
  count() %>% 
  ungroup() %>% 
  mutate(per=`n`/sum(`n`)) %>% 
  arrange(desc(a))
data$label <- scales::percent(data$per)
ggplot(data=data)+
  geom_bar(aes(x="", y=per, fill=a), stat="identity", width = 1)+
  coord_polar("y", start=0)+
  theme_void()+
  geom_text(aes(x=1, y = cumsum(per) - per/2, label=label)) +
  labs(title = "Attrition Rates: Job Level = 4",fill = "Attrition")

#Plot Pie chart for Job Level = 5
data <- data.frame(a=Level5$Attrition)
data <- data %>% 
  group_by(a) %>% 
  count() %>% 
  ungroup() %>% 
  mutate(per=`n`/sum(`n`)) %>% 
  arrange(desc(a))
data$label <- scales::percent(data$per)
ggplot(data=data)+
  geom_bar(aes(x="", y=per, fill=a), stat="identity", width = 1)+
  coord_polar("y", start=0)+
  theme_void()+
  geom_text(aes(x=1, y = cumsum(per) - per/2, label=label)) +
  labs(title = "Attrition Rates: Job Level = 5",fill = "Attrition")
```

##JobRole_OT Data
```{r JobRole_OT}
#Split into Job ROle Subsets for Plotting
Healthcare = CaseStudy %>% filter(JobRole == "Healthcare Representative")
HRRole = CaseStudy %>% filter(JobRole == "Human Resources")
Lab = CaseStudy %>% filter(JobRole == "Laboratory Technician")
Manager = CaseStudy %>% filter(JobRole == "Manager")
Manu = CaseStudy %>% filter(JobRole == "Manufacturing Director")
ResearchDirector = CaseStudy %>% filter(JobRole == "Research Director")
ResearchScientist = CaseStudy %>% filter(JobRole == "Research Scientist")
SalesExec = CaseStudy %>% filter(JobRole == "Sales Executive")
SalesRep = CaseStudy %>% filter(JobRole == "Sales Representative")

#Split into Overtime Status Subsets for Plotting
NOOT = CaseStudy %>% filter(OverTime == "No")
OT = CaseStudy %>% filter(OverTime == "Yes")

#Split into Job ROle & OT Status Subsets for Plotting
HealthcareOT = CaseStudy %>% filter(JobRole_OT == "Healthcare Representative_Yes")
HealthcareNOOT = CaseStudy %>% filter(JobRole_OT == "Healthcare Representative_No")
HRRoleOT = CaseStudy %>% filter(JobRole_OT == "Human Resources_Yes")
HRRoleNOOT = CaseStudy %>% filter(JobRole_OT == "Human Resources_No")
LabOT = CaseStudy %>% filter(JobRole_OT == "Laboratory Technician_Yes")
LabNOOT = CaseStudy %>% filter(JobRole_OT == "Laboratory Technician_No")
ManagerOT = CaseStudy %>% filter(JobRole_OT == "Manager_Yes")
ManagerNOOT = CaseStudy %>% filter(JobRole_OT == "Manager_No")
ManuOT = CaseStudy %>% filter(JobRole_OT == "Manufacturing Director_Yes")
ManuNOOT = CaseStudy %>% filter(JobRole_OT == "Manufacturing Director_No")
ResearchDirectorOT = CaseStudy %>% filter(JobRole_OT == "Research Director_Yes")
ResearchDirectorNOOT = CaseStudy %>% filter(JobRole_OT == "Research Director_No")
ResearchScientistOT = CaseStudy %>% filter(JobRole_OT == "Research Scientist_Yes")
ResearchScientistNOOT = CaseStudy %>% filter(JobRole_OT == "Research Scientist_No")
SalesExecOT = CaseStudy %>% filter(JobRole_OT == "Sales Executive_Yes")
SalesExecNOOT = CaseStudy %>% filter(JobRole_OT == "Sales Executive_No")
SalesRepOT = CaseStudy %>% filter(JobRole_OT == "Sales Representative_Yes")
SalesRepNOOT = CaseStudy %>% filter(JobRole_OT == "Sales Representative_No")

#Plot Pie charts for Healthcare 
#OT
data <- data.frame(a=HealthcareOT$Attrition)
data <- data %>% 
  group_by(a) %>% 
  count() %>% 
  ungroup() %>% 
  mutate(per=`n`/sum(`n`)) %>% 
  arrange(desc(a))
data$label <- scales::percent(data$per)
ggplot(data=data)+
  geom_bar(aes(x="", y=per, fill=a), stat="identity", width = 1)+
  coord_polar("y", start=0)+
  theme_void()+
  geom_text(aes(x=1, y = cumsum(per) - per/2, label=label)) +
  labs(title = "Attrition Rates: Overtime Healthcare",fill = "Attrition")

#No OT
data <- data.frame(a=HealthcareNOOT$Attrition)
data <- data %>% 
  group_by(a) %>% 
  count() %>% 
  ungroup() %>% 
  mutate(per=`n`/sum(`n`)) %>% 
  arrange(desc(a))
data$label <- scales::percent(data$per)
ggplot(data=data)+
  geom_bar(aes(x="", y=per, fill=a), stat="identity", width = 1)+
  coord_polar("y", start=0)+
  theme_void()+
  geom_text(aes(x=1, y = cumsum(per) - per/2, label=label)) +
  labs(title = "Attrition Rates: Non-Overtime Healthcare",fill = "Attrition")

#Plot Pie charts for HR 
#OT
data <- data.frame(a=HRRoleOT$Attrition)
data <- data %>% 
  group_by(a) %>% 
  count() %>% 
  ungroup() %>% 
  mutate(per=`n`/sum(`n`)) %>% 
  arrange(desc(a))
data$label <- scales::percent(data$per)
ggplot(data=data)+
  geom_bar(aes(x="", y=per, fill=a), stat="identity", width = 1)+
  coord_polar("y", start=0)+
  theme_void()+
  geom_text(aes(x=1, y = cumsum(per) - per/2, label=label)) +
  labs(title = "Attrition Rates: Overtime Human Resources",fill = "Attrition")

#No OT
data <- data.frame(a=HRRoleNOOT$Attrition)
data <- data %>% 
  group_by(a) %>% 
  count() %>% 
  ungroup() %>% 
  mutate(per=`n`/sum(`n`)) %>% 
  arrange(desc(a))
data$label <- scales::percent(data$per)
ggplot(data=data)+
  geom_bar(aes(x="", y=per, fill=a), stat="identity", width = 1)+
  coord_polar("y", start=0)+
  theme_void()+
  geom_text(aes(x=1, y = cumsum(per) - per/2, label=label)) +
  labs(title = "Attrition Rates: Non-Overtime Human Resources",fill = "Attrition")

#Plot Pie charts for Lab Tech 
#OT
data <- data.frame(a=LabOT$Attrition)
data <- data %>% 
  group_by(a) %>% 
  count() %>% 
  ungroup() %>% 
  mutate(per=`n`/sum(`n`)) %>% 
  arrange(desc(a))
data$label <- scales::percent(data$per)
ggplot(data=data)+
  geom_bar(aes(x="", y=per, fill=a), stat="identity", width = 1)+
  coord_polar("y", start=0)+
  theme_void()+
  geom_text(aes(x=1, y = cumsum(per) - per/2, label=label)) +
  labs(title = "Attrition Rates: Overtime Lab Tech",fill = "Attrition")

#No OT
data <- data.frame(a=LabNOOT$Attrition)
data <- data %>% 
  group_by(a) %>% 
  count() %>% 
  ungroup() %>% 
  mutate(per=`n`/sum(`n`)) %>% 
  arrange(desc(a))
data$label <- scales::percent(data$per)
ggplot(data=data)+
  geom_bar(aes(x="", y=per, fill=a), stat="identity", width = 1)+
  coord_polar("y", start=0)+
  theme_void()+
  geom_text(aes(x=1, y = cumsum(per) - per/2, label=label)) +
  labs(title = "Attrition Rates: Non-Overtime Lab Tech",fill = "Attrition")

#Plot Pie charts for Manager 
#OT
data <- data.frame(a=ManagerOT$Attrition)
data <- data %>% 
  group_by(a) %>% 
  count() %>% 
  ungroup() %>% 
  mutate(per=`n`/sum(`n`)) %>% 
  arrange(desc(a))
data$label <- scales::percent(data$per)
ggplot(data=data)+
  geom_bar(aes(x="", y=per, fill=a), stat="identity", width = 1)+
  coord_polar("y", start=0)+
  theme_void()+
  geom_text(aes(x=1, y = cumsum(per) - per/2, label=label)) +
  labs(title = "Attrition Rates: Overtime Manager",fill = "Attrition")

#No OT
data <- data.frame(a=ManagerNOOT$Attrition)
data <- data %>% 
  group_by(a) %>% 
  count() %>% 
  ungroup() %>% 
  mutate(per=`n`/sum(`n`)) %>% 
  arrange(desc(a))
data$label <- scales::percent(data$per)
ggplot(data=data)+
  geom_bar(aes(x="", y=per, fill=a), stat="identity", width = 1)+
  coord_polar("y", start=0)+
  theme_void()+
  geom_text(aes(x=1, y = cumsum(per) - per/2, label=label)) +
  labs(title = "Attrition Rates: Non-Overtime Manager",fill = "Attrition")


#Plot Pie charts for Manufacturing Director 
#OT
data <- data.frame(a=ManuOT$Attrition)
data <- data %>% 
  group_by(a) %>% 
  count() %>% 
  ungroup() %>% 
  mutate(per=`n`/sum(`n`)) %>% 
  arrange(desc(a))
data$label <- scales::percent(data$per)
ggplot(data=data)+
  geom_bar(aes(x="", y=per, fill=a), stat="identity", width = 1)+
  coord_polar("y", start=0)+
  theme_void()+
  geom_text(aes(x=1, y = cumsum(per) - per/2, label=label)) +
  labs(title = "Attrition Rates: Overtime Manufacturing Director",fill = "Attrition")

#No OT
data <- data.frame(a=ManuNOOT$Attrition)
data <- data %>% 
  group_by(a) %>% 
  count() %>% 
  ungroup() %>% 
  mutate(per=`n`/sum(`n`)) %>% 
  arrange(desc(a))
data$label <- scales::percent(data$per)
ggplot(data=data)+
  geom_bar(aes(x="", y=per, fill=a), stat="identity", width = 1)+
  coord_polar("y", start=0)+
  theme_void()+
  geom_text(aes(x=1, y = cumsum(per) - per/2, label=label)) +
  labs(title = "Attrition Rates: Non-Overtime Manufacturing Director",fill = "Attrition")

#Plot Pie charts for Research Director 
#OT
data <- data.frame(a=ResearchDirectorOT$Attrition)
data <- data %>% 
  group_by(a) %>% 
  count() %>% 
  ungroup() %>% 
  mutate(per=`n`/sum(`n`)) %>% 
  arrange(desc(a))
data$label <- scales::percent(data$per)
ggplot(data=data)+
  geom_bar(aes(x="", y=per, fill=a), stat="identity", width = 1)+
  coord_polar("y", start=0)+
  theme_void()+
  geom_text(aes(x=1, y = cumsum(per) - per/2, label=label)) +
  labs(title = "Attrition Rates: Overtime Research Director",fill = "Attrition")

#No OT
data <- data.frame(a=ResearchDirectorNOOT$Attrition)
data <- data %>% 
  group_by(a) %>% 
  count() %>% 
  ungroup() %>% 
  mutate(per=`n`/sum(`n`)) %>% 
  arrange(desc(a))
data$label <- scales::percent(data$per)
ggplot(data=data)+
  geom_bar(aes(x="", y=per, fill=a), stat="identity", width = 1)+
  coord_polar("y", start=0)+
  theme_void()+
  geom_text(aes(x=1, y = cumsum(per) - per/2, label=label)) +
  labs(title = "Attrition Rates: Non-Overtime Research Director",fill = "Attrition")

#Plot Pie charts for Research Scientist
#OT
data <- data.frame(a=ResearchScientistOT$Attrition)
data <- data %>% 
  group_by(a) %>% 
  count() %>% 
  ungroup() %>% 
  mutate(per=`n`/sum(`n`)) %>% 
  arrange(desc(a))
data$label <- scales::percent(data$per)
ggplot(data=data)+
  geom_bar(aes(x="", y=per, fill=a), stat="identity", width = 1)+
  coord_polar("y", start=0)+
  theme_void()+
  geom_text(aes(x=1, y = cumsum(per) - per/2, label=label)) +
  labs(title = "Attrition Rates: Overtime Research Scientist",fill = "Attrition")

#No OT
data <- data.frame(a=ResearchScientistNOOT$Attrition)
data <- data %>% 
  group_by(a) %>% 
  count() %>% 
  ungroup() %>% 
  mutate(per=`n`/sum(`n`)) %>% 
  arrange(desc(a))
data$label <- scales::percent(data$per)
ggplot(data=data)+
  geom_bar(aes(x="", y=per, fill=a), stat="identity", width = 1)+
  coord_polar("y", start=0)+
  theme_void()+
  geom_text(aes(x=1, y = cumsum(per) - per/2, label=label)) +
  labs(title = "Attrition Rates: Non-Overtime Research Scientist",fill = "Attrition")

#Plot Pie charts for Sales Executive
#OT
data <- data.frame(a=SalesExecOT$Attrition)
data <- data %>% 
  group_by(a) %>% 
  count() %>% 
  ungroup() %>% 
  mutate(per=`n`/sum(`n`)) %>% 
  arrange(desc(a))
data$label <- scales::percent(data$per)
ggplot(data=data)+
  geom_bar(aes(x="", y=per, fill=a), stat="identity", width = 1)+
  coord_polar("y", start=0)+
  theme_void()+
  geom_text(aes(x=1, y = cumsum(per) - per/2, label=label)) +
  labs(title = "Attrition Rates: Overtime Sales Executive",fill = "Attrition")

#No OT
data <- data.frame(a=SalesExecNOOT$Attrition)
data <- data %>% 
  group_by(a) %>% 
  count() %>% 
  ungroup() %>% 
  mutate(per=`n`/sum(`n`)) %>% 
  arrange(desc(a))
data$label <- scales::percent(data$per)
ggplot(data=data)+
  geom_bar(aes(x="", y=per, fill=a), stat="identity", width = 1)+
  coord_polar("y", start=0)+
  theme_void()+
  geom_text(aes(x=1, y = cumsum(per) - per/2, label=label)) +
  labs(title = "Attrition Rates: Non-Overtime Sales Executive",fill = "Attrition")

#Plot Pie charts for Sales Rep
#OT
data <- data.frame(a=SalesRepOT$Attrition)
data <- data %>% 
  group_by(a) %>% 
  count() %>% 
  ungroup() %>% 
  mutate(per=`n`/sum(`n`)) %>% 
  arrange(desc(a))
data$label <- scales::percent(data$per)
ggplot(data=data)+
  geom_bar(aes(x="", y=per, fill=a), stat="identity", width = 1)+
  coord_polar("y", start=0)+
  theme_void()+
  geom_text(aes(x=1, y = cumsum(per) - per/2, label=label)) +
  labs(title = "Attrition Rates: Overtime Sales Rep",fill = "Attrition")

#No OT
data <- data.frame(a=SalesRepNOOT$Attrition)
data <- data %>% 
  group_by(a) %>% 
  count() %>% 
  ungroup() %>% 
  mutate(per=`n`/sum(`n`)) %>% 
  arrange(desc(a))
data$label <- scales::percent(data$per)
ggplot(data=data)+
  geom_bar(aes(x="", y=per, fill=a), stat="identity", width = 1)+
  coord_polar("y", start=0)+
  theme_void()+
  geom_text(aes(x=1, y = cumsum(per) - per/2, label=label)) +
  labs(title = "Attrition Rates: Non-Overtime Sales Executive",fill = "Attrition")
```

#Split into Stock Option Subsets for Plotting
```{r Stock Option Levels}
Options0 = CaseStudy %>% filter(StockOptionLevel == "0")
Options1 = CaseStudy %>% filter(StockOptionLevel == "1")
Options2 = CaseStudy %>% filter(StockOptionLevel == "2")
Options3 = CaseStudy %>% filter(StockOptionLevel == "3")

#Plot Pie chart for Stock Option Level 0
data <- data.frame(a=Options0$Attrition)
data <- data %>% 
  group_by(a) %>% 
  count() %>% 
  ungroup() %>% 
  mutate(per=`n`/sum(`n`)) %>% 
  arrange(desc(a))
data$label <- scales::percent(data$per)
ggplot(data=data)+
  geom_bar(aes(x="", y=per, fill=a), stat="identity", width = 1)+
  coord_polar("y", start=0)+
  theme_void()+
  geom_text(aes(x=1, y = cumsum(per) - per/2, label=label)) +
  labs(title = "Attrition Rates: Stock Option Level = 0",fill = "Attrition")

#Plot Pie chart for Stock Option Level 1
data <- data.frame(a=Options1$Attrition)
data <- data %>% 
  group_by(a) %>% 
  count() %>% 
  ungroup() %>% 
  mutate(per=`n`/sum(`n`)) %>% 
  arrange(desc(a))
data$label <- scales::percent(data$per)
ggplot(data=data)+
  geom_bar(aes(x="", y=per, fill=a), stat="identity", width = 1)+
  coord_polar("y", start=0)+
  theme_void()+
  geom_text(aes(x=1, y = cumsum(per) - per/2, label=label)) +
  labs(title = "Attrition Rates: Stock Option Level = 1",fill = "Attrition")

#Plot Pie chart for Stock Option Level 2
data <- data.frame(a=Options2$Attrition)
data <- data %>% 
  group_by(a) %>% 
  count() %>% 
  ungroup() %>% 
  mutate(per=`n`/sum(`n`)) %>% 
  arrange(desc(a))
data$label <- scales::percent(data$per)
ggplot(data=data)+
  geom_bar(aes(x="", y=per, fill=a), stat="identity", width = 1)+
  coord_polar("y", start=0)+
  theme_void()+
  geom_text(aes(x=1, y = cumsum(per) - per/2, label=label)) +
  labs(title = "Attrition Rates: Stock Option Level = 2",fill = "Attrition")

#Plot Pie chart for Stock Option Level 3
data <- data.frame(a=Options3$Attrition)
data <- data %>% 
  group_by(a) %>% 
  count() %>% 
  ungroup() %>% 
  mutate(per=`n`/sum(`n`)) %>% 
  arrange(desc(a))
data$label <- scales::percent(data$per)
ggplot(data=data)+
  geom_bar(aes(x="", y=per, fill=a), stat="identity", width = 1)+
  coord_polar("y", start=0)+
  theme_void()+
  geom_text(aes(x=1, y = cumsum(per) - per/2, label=label)) +
  labs(title = "Attrition Rates: Stock Option Level = 3",fill = "Attrition")
```

```{r Marital Status}
##Marital Status
#Split into Marital Status Subsets for Plotting
Divorced = CaseStudy %>% filter(MaritalStatus == "Divorced")
Married = CaseStudy %>% filter(MaritalStatus == "Married")
Single = CaseStudy %>% filter(MaritalStatus == "Single")

#Plot Pie chart for Divorced
data <- data.frame(a=Divorced$Attrition)
data <- data %>% 
  group_by(a) %>% 
  count() %>% 
  ungroup() %>% 
  mutate(per=`n`/sum(`n`)) %>% 
  arrange(desc(a))
data$label <- scales::percent(data$per)
ggplot(data=data)+
  geom_bar(aes(x="", y=per, fill=a), stat="identity", width = 1)+
  coord_polar("y", start=0)+
  theme_void()+
  geom_text(aes(x=1, y = cumsum(per) - per/2, label=label)) +
  labs(title = "Attrition Rates: Divorced",fill = "Attrition")

#Plot Pie chart for Married
data <- data.frame(a=Married$Attrition)
data <- data %>% 
  group_by(a) %>% 
  count() %>% 
  ungroup() %>% 
  mutate(per=`n`/sum(`n`)) %>% 
  arrange(desc(a))
data$label <- scales::percent(data$per)
ggplot(data=data)+
  geom_bar(aes(x="", y=per, fill=a), stat="identity", width = 1)+
  coord_polar("y", start=0)+
  theme_void()+
  geom_text(aes(x=1, y = cumsum(per) - per/2, label=label)) +
  labs(title = "Attrition Rates: Married",fill = "Attrition")

#Plot Pie chart for Married
data <- data.frame(a=Single$Attrition)
data <- data %>% 
  group_by(a) %>% 
  count() %>% 
  ungroup() %>% 
  mutate(per=`n`/sum(`n`)) %>% 
  arrange(desc(a))
data$label <- scales::percent(data$per)
ggplot(data=data)+
  geom_bar(aes(x="", y=per, fill=a), stat="identity", width = 1)+
  coord_polar("y", start=0)+
  theme_void()+
  geom_text(aes(x=1, y = cumsum(per) - per/2, label=label)) +
  labs(title = "Attrition Rates: Single",fill = "Attrition")
```

#Split into Work/Life Balance Subsets for Plotting
```{r WorkLife Balance}
WorkLife1 = CaseStudy %>% filter(WorkLifeBalance == 1)
WorkLife2 = CaseStudy %>% filter(WorkLifeBalance == 2)
WorkLife3 = CaseStudy %>% filter(WorkLifeBalance == 3)
WorkLife4 = CaseStudy %>% filter(WorkLifeBalance == 4)

#Plot Pie chart for Work Life Balance = 1
data <- data.frame(a=WorkLife1$Attrition)
data <- data %>% 
  group_by(a) %>% 
  count() %>% 
  ungroup() %>% 
  mutate(per=`n`/sum(`n`)) %>% 
  arrange(desc(a))
data$label <- scales::percent(data$per)
ggplot(data=data)+
  geom_bar(aes(x="", y=per, fill=a), stat="identity", width = 1)+
  coord_polar("y", start=0)+
  theme_void()+
  geom_text(aes(x=1, y = cumsum(per) - per/2, label=label)) +
  labs(title = "Attrition Rates: Work/Life Balance = 1",fill = "Attrition")

#Plot Pie chart for Work Life Balance = 2
data <- data.frame(a=WorkLife2$Attrition)
data <- data %>% 
  group_by(a) %>% 
  count() %>% 
  ungroup() %>% 
  mutate(per=`n`/sum(`n`)) %>% 
  arrange(desc(a))
data$label <- scales::percent(data$per)
ggplot(data=data)+
  geom_bar(aes(x="", y=per, fill=a), stat="identity", width = 1)+
  coord_polar("y", start=0)+
  theme_void()+
  geom_text(aes(x=1, y = cumsum(per) - per/2, label=label)) +
  labs(title = "Attrition Rates: Work/Life Balance = 2",fill = "Attrition")

#Plot Pie chart for Work Life Balance = 3
data <- data.frame(a=WorkLife3$Attrition)
data <- data %>% 
  group_by(a) %>% 
  count() %>% 
  ungroup() %>% 
  mutate(per=`n`/sum(`n`)) %>% 
  arrange(desc(a))
data$label <- scales::percent(data$per)
ggplot(data=data)+
  geom_bar(aes(x="", y=per, fill=a), stat="identity", width = 1)+
  coord_polar("y", start=0)+
  theme_void()+
  geom_text(aes(x=1, y = cumsum(per) - per/2, label=label)) +
  labs(title = "Attrition Rates: Work/Life Balance = 3",fill = "Attrition")

#Plot Pie chart for Work Life Balance = 4
data <- data.frame(a=WorkLife4$Attrition)
data <- data %>% 
  group_by(a) %>% 
  count() %>% 
  ungroup() %>% 
  mutate(per=`n`/sum(`n`)) %>% 
  arrange(desc(a))
data$label <- scales::percent(data$per)
ggplot(data=data)+
  geom_bar(aes(x="", y=per, fill=a), stat="identity", width = 1)+
  coord_polar("y", start=0)+
  theme_void()+
  geom_text(aes(x=1, y = cumsum(per) - per/2, label=label)) +
  labs(title = "Attrition Rates: Work/Life Balance = 4",fill = "Attrition")
```

#Split into Gender Subsets for Plotting
```{r Gender}
Male = CaseStudy %>% filter(Gender == "Male")
Female = CaseStudy %>% filter(Gender == "Female")

#Plot Pie chart for Men
data <- data.frame(a=Male$Attrition)
data <- data %>% 
  group_by(a) %>% 
  count() %>% 
  ungroup() %>% 
  mutate(per=`n`/sum(`n`)) %>% 
  arrange(desc(a))
data$label <- scales::percent(data$per)
ggplot(data=data)+
  geom_bar(aes(x="", y=per, fill=a), stat="identity", width = 1)+
  coord_polar("y", start=0)+
  theme_void()+
  geom_text(aes(x=1, y = cumsum(per) - per/2, label=label)) +
  labs(title = "Attrition Rates: Men",fill = "Attrition")

#Plot Pie chart for Women
data <- data.frame(a=Female$Attrition)
data <- data %>% 
  group_by(a) %>% 
  count() %>% 
  ungroup() %>% 
  mutate(per=`n`/sum(`n`)) %>% 
  arrange(desc(a))
data$label <- scales::percent(data$per)
ggplot(data=data)+
  geom_bar(aes(x="", y=per, fill=a), stat="identity", width = 1)+
  coord_polar("y", start=0)+
  theme_void()+
  geom_text(aes(x=1, y = cumsum(per) - per/2, label=label)) +
  labs(title = "Attrition Rates: Women",fill = "Attrition")
```

# Monthly Income Boxplot
```{r Monthly Income}
CaseStudy %>% 
  ggplot(aes(y = MonthlyIncome,fill = Attrition)) + 
  geom_boxplot() + 
  facet_wrap(~Attrition) +
  labs(title = "Attrition by Monthly Income",y = "Monthly Income")
```

# Total Working Years
```{r Total Working Years}
CaseStudy %>% ggplot(aes(y = TotalWorkingYears,fill = Attrition)) + 
  geom_boxplot() + 
  facet_wrap(~Attrition) +
  labs(title = "Attrition by Total Working Years",y = "Total Working Years")
```



##KNN
```{r kknn}
testmodel = train.kknn(Attrition ~ JobRole_OT + MonthlyIncomeHL + JobInvolvement + 
                         WorkLifeBalance + TotalWorkHL + JobLevel + StockOptionLevel + 
                         MaritalStatus + Gender,
                       data = CaseStudy,kmax = 15,
                       kernel = c("rectangular","triangular","epanechnikov","gaussian","rank","optimal"),scale = TRUE)
model = train.kknn(Attrition ~ JobRole_OT + MonthlyIncomeHL + JobInvolvement + 
                     WorkLifeBalance + TotalWorkHL + JobLevel + StockOptionLevel + 
                     MaritalStatus + Gender,
                   data = CaseStudy,ks = testmodel$best.parameters$k,
                   kernel = testmodel$best.parameters$kernel,scale = TRUE)
preds = predict(model,newdata = CaseStudy)
CM = confusionMatrix(preds,CaseStudy$Attrition)
CM
```

##Set up the data for the competition set
```{r Set up No Attrition Date}
NoAttrition[,names] <- lapply(NoAttrition[,names],factor,ordered = TRUE)
#Create new variables for job OT & Job Gender
NoAttrition <- NoAttrition %>% mutate(JobRole_OT = paste(JobRole,OverTime,sep = "_"))
NoAttrition$JobRole_OT = as.factor(NoAttrition$JobRole_OT)

NoAttrition$MonthlyIncomeHL = as.factor(if_else(NoAttrition$MonthlyIncome <= IncomeSum[2],"Lowest",
                                               if_else(NoAttrition$MonthlyIncome <= IncomeSum[3],"Medium/Low",
                                                       if_else(NoAttrition$MonthlyIncome <= IncomeSum[4],"Medium/High","High"))))

NoAttrition$TotalWorkHL = as.factor(if_else(NoAttrition$TotalWorkingYears <= TotalWorkSum[2],"<=6",
                                           if_else(NoAttrition$TotalWorkingYears <= TotalWorkSum[3],"6< <=10",
                                                   if_else(NoAttrition$TotalWorkingYears <= TotalWorkSum[5],"10< <=15",">15"))))
```

## Run predictions on the competition set using kknn
```{r kknn on competition set}
testmodel = train.kknn(Attrition ~ JobRole_OT + MonthlyIncomeHL + JobInvolvement + 
                         WorkLifeBalance + TotalWorkHL + JobLevel + StockOptionLevel + 
                         MaritalStatus + Gender,
                       data = CaseStudy,kmax = 25,
                       kernel = c("rectangular","triangular","epanechnikov","gaussian","rank","optimal"),scale = TRUE)
model = train.kknn(Attrition ~ JobRole_OT + MonthlyIncomeHL + JobInvolvement + 
                     WorkLifeBalance + TotalWorkHL + JobLevel + StockOptionLevel + 
                     MaritalStatus + Gender,
                   data = CaseStudy,ks = testmodel$best.parameters$k,
                   kernel = testmodel$best.parameters$kernel,scale = TRUE)
NoAttrition$predictions = predict(model,newdata = NoAttrition)
```

##Write the data to a csv for submission
```{r write csv}
NoAttritionSubmit = NoAttrition[,c("ID","predictions")]
path_out = "C:\\Users\\Steve Mann\\OneDrive\\Desktop\\"
write.csv(NoAttritionSubmit,"C:\\Users\\Steve Mann\\OneDrive\\Desktop\\Case2PredictionsMann_Attrition.csv")
```

##Write a model to predict Monthly Income.
##Model was adjusted manually by stepwise regression. It is likely there are other variables
##or ineraction variables that would be considered significant, however, as the mean RMSE on 
##1000 iterations of 70/30 train/test split is ~1000, I decided to stop.
```{r MLR, include=FALSE}
CaseStudyRegression <- glm(formula = CaseStudy$MonthlyIncome ~ CaseStudy$OverTime + CaseStudy$TotalWorkingYears + CaseStudy$JobRole_OT + CaseStudy$JobLevel,data = CaseStudy)
summary(CaseStudyRegression)

CaseStudy$PredictedMonthly = predict(CaseStudyRegression)
rmse(CaseStudy$MonthlyIncome,CaseStudy$PredictedMonthly)
rsq(CaseStudyRegression,adj = TRUE)

##Loop to ensure good RMSE & no overfit
iterations = 500
RootMSE = matrix(nrow = iterations)
R_Square = matrix(nrow = iterations)
for(j in 1:iterations)
{
  
  data <- CaseStudy[,c("MonthlyIncome","OverTime","TotalWorkingYears","JobRole_OT","JobLevel")]
  Training_obs <- data$MonthlyIncome %>% createDataPartition(p = 0.7,list = FALSE)
  train <- data[Training_obs,]
  test <- data[-Training_obs,]
  model <- lm(MonthlyIncome ~ .,data = train)
  test$predictions <- model %>% predict(test)
  RootMSE[j] = rmse(test$predictions,test$MonthlyIncome)
  R_Square[j] = R2(test$predictions,test$MonthlyIncome)
}
```

#Summary statistics of model
```{r Model Summary}
summary(model)
MeanRMSE = colMeans(RootMSE)
MeanR_Square = colMeans(R_Square)
MeanRMSE
MeanR_Square
```

```{r Salary Predictions for Submission}
NoSalary
newdata <- NoSalary[,c("OverTime","TotalWorkingYears","JobRole_OT","JobLevel")]
model <- lm(MonthlyIncome ~ .,data = data)
newdata$predictions <- model %>% predict(newdata)
newdata$ID = NoSalary$ID
newdata_submit = newdata[,c("ID","predictions")]
path_out = "C:\\Users\\Steve Mann\\OneDrive\\Desktop\\"
write.csv(newdata_submit,"C:\\Users\\Steve Mann\\OneDrive\\Desktop\\Case2PredictionsMann_Salary.csv")
newdata_submit
```

