---
title: "Global Life Expectancy Prediction Using Random Forest"
author: "Steven Mann"
date: "9/28/2020"
output:
  html_document: default
  word_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r Read in Life Expectancy Data from Kaggle}
#Life Expectancy Data from Kaggle
Life = read.csv(file.choose(),header = TRUE)
```

```{r Load Libraries}
library(Amelia)
library(tidyverse)
library(DataCombine)
```

```{r See what data is missing}
missmap(Life)
```

#We see Population is biggest missing value. See if we can find that data on UN website.
#Population Data downloaded from the UN
#URL https://population.un.org/wpp/Download/Standard/Population/

```{r Read in Demographic Data from UN Database}
UNPOP = read.csv(file.choose(),header=TRUE)
```

```{r Filter and Select Columns for imputing}
#filter UN data set down to size
#Grab only columns of interest
UNPOP2 = data.frame(Country = UNPOP$Location,Year = UNPOP$Time,Population = UNPOP$PopTotal*1000)
#Filter to only time frame of interest
UNPOPclean = UNPOP2 %>% filter(Year >= 2000 & Year <= 2015)
```

#Initially planned to replace only N/A values from original dataset. However, an inspection of the dataset
#reveals that many population factors clearly aren't correct. An inspection of just the first 2 countries 
#in the list, Afghanistan and Albania, reveal that year-over-year popluation fluctuations are by multiples or
#divisions of 10x or 100x. Decided instead to replace the entire population data with data obtained from the 
#UN population database.

```{r Check if any missing values in the UN Data}
#No Missing Values
missmap(UNPOPclean)
```

#Some country names do not match our other data set. This will mess with the merge.
#Rename countries to fit with our dataset Using FindReplace function

```{r Merging in Complete Population Data}
#Create Data frame to replace country names
Replace = data.frame(from = c("Dem. People's Republic of Korea","Micronesia \\(Fed. States of\\)",
                              "Eswatini","North Macedonia","United Kingdom"), 
                     to = c("Democratic People's Republic of Korea","Micronesia \\(Federated States of\\)",
                            "Swaziland","The former Yugoslav republic of Macedonia",
                            "United Kingdom of Great Britain and Northern Ireland"))

#Replace Country Names
UNPOPclean2 = FindReplace(data = UNPOPclean,Var = "Country",replaceData = Replace,
                          from = "from",to = "to",exact = FALSE)

#Merge Clean Population Data into our dataset
Lifeclean = left_join(Life, UNPOPclean2, by = c("Country", "Year"))

#Delete old Population Column
Lifeclean2 = Lifeclean[,-19]

#Rename Column
Lifeclean3 = Lifeclean2 %>% rename("Population" = Population.y)

#Filter out N/A Life Expectancies
Lifeclean3 = Lifeclean3 %>% filter(is.na(Life.expectancy) == "FALSE")

#Sanity Check - All good!
missmap(Lifeclean3)
```


```{r Merging in Complete GDP Data}
##Same issues with GDP 
##Get data from UN
## URL http://data.un.org/Data.aspx?d=SNAAMA&f=grID%3a101%3bcurrID%3aUSD%3bpcFlag%3a0%3bitID%3a9
##filtered on the front-end to only include 2000 - 2015

GDP = read.csv(file.choose(),header = TRUE)

#Rename Columns
GDP2 = data.frame(Country = GDP$Country.or.Area,Year = GDP$Year,GDP = GDP$Value)

#Some country names do not match our other data set. This will mess with the merge.
#Rename countries to fit with our dataset Using FindReplace function

#Create Data frame to replace country names
Replace2 = data.frame(from = c("China, People's Republic of",
                               "Iran, Islamic Republic of","Kingdom of Eswatini",
                               "Republic of North Macedonia","United States",
                               "United Republic of Tanzania: Mainland","Former Sudan"), 
                      to = c("China","Iran \\(Islamic Republic of\\)","Swaziland",
                             "The former Yugoslav republic of Macedonia","United States of America","United Republic of Tanzania","Sudan"))

#Replace Country Names
GDP3 = FindReplace(data = GDP2,Var = "Country",replaceData = Replace2,
                   from = "from",to = "to",exact = FALSE)

#Join Dataset with new GDP data
Lifeclean4 = left_join(Lifeclean3, GDP3, by = c("Country", "Year"))

#Delete old GDP Column
Lifeclean5 = Lifeclean4[,-18]

#Rename Column
Lifeclean6 = Lifeclean5 %>% rename("GDP" = GDP.y)

#Remaining N/A's are Niue, & South Sudan. South Sudan didn't exist prior to 2008. Niue only has 1 year of data in original dataset. Reasonable.
Lifeclean6 %>% filter(is.na(GDP)==TRUE)

#Sanity checks
summary(Lifeclean6$Population)
summary(Lifeclean6$GDP)
```

```{r Impute other missing values}
#See what else is missing
missmap(Lifeclean6)

##Compute Median Hepatitis B by Country
HepBCountry = Lifeclean6 %>% 
  filter(is.na(Hepatitis.B) == "FALSE") %>% 
  group_by(Country) %>% 
  summarize(MedianHBCountry = median(Hepatitis.B))

##Compute Median Hepatitis B by Continent
HepBContinent = Lifeclean6 %>% 
  filter(is.na(Hepatitis.B) == "FALSE") %>%
  group_by(Continent) %>%
  summarize(MedianHBContinent = median(Hepatitis.B))

#Merge Medians by Continent
Lifeclean7 <- merge(x = Lifeclean6,y = HepBContinent,by = "Continent",all.x = TRUE)

#Merge Medians by Country
Lifeclean8 <- merge(x = Lifeclean7,y = HepBCountry,by = "Country",all.x = TRUE)

#If statement to cascade which median to use: Non-N/A Value > Country Median
Lifeclean9 = within(Lifeclean8, 
       HepB1 <- 
         ifelse(is.na(Hepatitis.B)=="TRUE",
                MedianHBCountry,
              Hepatitis.B))

#If statement to cascade which median to use: Non-N/A value > Country Median > Continent Median
## Could be more efficient, but couldn't get nested ifelse to work
Lifeclean10 = within(Lifeclean9, 
                    HepBFinal <- 
                      ifelse(is.na(HepB1)=="TRUE",
                             MedianHBContinent,
                             HepB1))

#Drop Unneeded Columns
Lifeclean11 = select(Lifeclean10,-c(Hepatitis.B,MedianHBContinent,MedianHBCountry,HepB1))

##SanityCheck
missmap(Lifeclean11)

##Compute Median by Country
ExpCountry = Lifeclean11 %>% 
  filter(is.na(Total.expenditure) == "FALSE") %>% 
  group_by(Country) %>% 
  summarize(MedianexpCountry = median(Total.expenditure))

##Compute Median by Continent
ExpContinent = Lifeclean11 %>% 
  filter(is.na(Total.expenditure) == "FALSE") %>%
  group_by(Continent) %>%
  summarize(MedianexpContinent = median(Total.expenditure))

#Merge Medians by Continent
Lifeclean12 <- merge(x = Lifeclean11,y = ExpContinent,by = "Continent",all.x = TRUE)

#Merge Medians by Country
Lifeclean13 <- merge(x = Lifeclean12,y = ExpCountry,by = "Country",all.x = TRUE)

#If statement to cascade which median to use: Non-N/A Value > Country Median
Lifeclean14 = within(Lifeclean13, 
                    Exp1 <- 
                      ifelse(is.na(Total.expenditure)=="TRUE",
                             MedianexpCountry,
                             Total.expenditure))

#If statement to cascade which median to use: Non-N/A value > Country Median > Continent Median
## Could be more efficient, but couldn't get nested ifelse to work
Lifeclean15 = within(Lifeclean14, 
                     Total.expenditureFinal <- 
                       ifelse(is.na(Exp1)=="TRUE",
                              MedianexpContinent,
                              Exp1))

#Drop Unneeded Columns
Lifeclean16 = select(Lifeclean15,-c(Total.expenditure,MedianexpContinent,MedianexpCountry,Exp1))

##SanityCheck
missmap(Lifeclean16)

##Compute Median by Country
AlcCountry = Lifeclean16 %>% 
  filter(is.na(Alcohol) == "FALSE") %>% 
  group_by(Country) %>% 
  summarize(MedianAlcCountry = median(Alcohol))

##Compute Median by Continent
AlcContinent = Lifeclean16 %>% 
  filter(is.na(Alcohol) == "FALSE") %>%
  group_by(Continent) %>%
  summarize(MedianAlcContinent = median(Alcohol))

#Merge Medians by Continent
Lifeclean17 <- merge(x = Lifeclean16,y = AlcContinent,by = "Continent",all.x = TRUE)

#Merge Medians by Country
Lifeclean18 <- merge(x = Lifeclean17,y = AlcCountry,by = "Country",all.x = TRUE)

#If statement to cascade which median to use: Non-N/A Value > Country Median
Lifeclean19 = within(Lifeclean18, 
                     Alc1 <- 
                       ifelse(is.na(Alcohol)=="TRUE",
                              MedianAlcCountry,
                              Alcohol))

#If statement to cascade which median to use: Non-N/A value > Country Median > Continent Median
## Could be more efficient, but couldn't get nested ifelse to work
Lifeclean20 = within(Lifeclean19, 
                     AlcoholFinal <- 
                       ifelse(is.na(Alc1)=="TRUE",
                              MedianAlcContinent,
                              Alc1))

#Drop Unneeded Columns
Lifeclean21 = select(Lifeclean20,-c(Alcohol,MedianAlcContinent,MedianAlcCountry,Alc1))

##SanityCheck
missmap(Lifeclean21)

##Compute Median by Country
IncCountry = Lifeclean21 %>% 
  filter(is.na(Income.composition.of.resources) == "FALSE") %>% 
  group_by(Country) %>% 
  summarize(MedianIncCountry = median(Income.composition.of.resources))

##Compute Median by Continent
IncContinent = Lifeclean21 %>% 
  filter(is.na(Income.composition.of.resources) == "FALSE") %>% 
  group_by(Continent) %>% 
  summarize(MedianIncContinent = median(Income.composition.of.resources))

#Merge Medians by Continent
Lifeclean22 <- merge(x = Lifeclean21,y = IncContinent,by = "Continent",all.x = TRUE)

#Merge Medians by Country
Lifeclean23 <- merge(x = Lifeclean22,y = IncCountry,by = "Country",all.x = TRUE)

#If statement to cascade which median to use: Non-N/A Value > Country Median
Lifeclean24 = within(Lifeclean23, 
                     Inc1 <- 
                       ifelse(is.na(Income.composition.of.resources)=="TRUE",
                              MedianIncCountry,
                              Income.composition.of.resources))

#If statement to cascade which median to use: Non-N/A value > Country Median > Continent Median
## Could be more efficient, but couldn't get nested ifelse to work
Lifeclean25 = within(Lifeclean24, 
                     IncomeFinal <- 
                       ifelse(is.na(Inc1)=="TRUE",
                              MedianIncContinent,
                              Inc1))

#Drop Unneeded Columns
Lifeclean26 = select(Lifeclean25,-c(Income.composition.of.resources,MedianIncContinent,MedianIncCountry,Inc1))

##SanityCheck
missmap(Lifeclean26)

##Compute Median by Country
SchoolCountry = Lifeclean26 %>% 
  filter(is.na(Schooling) == "FALSE") %>% 
  group_by(Country) %>% 
  summarize(MedianSchoolCountry = median(Schooling))

##Compute Median by Continent
SchoolContinent = Lifeclean26 %>% 
  filter(is.na(Schooling) == "FALSE") %>% 
  group_by(Continent) %>% 
  summarize(MedianSchoolContinent = median(Schooling))

#Merge Medians by Continent
Lifeclean27 <- merge(x = Lifeclean26,y = SchoolContinent,by = "Continent",all.x = TRUE)

#Merge Medians by Country
Lifeclean28 <- merge(x = Lifeclean27,y = SchoolCountry,by = "Country",all.x = TRUE)

#If statement to cascade which median to use: Non-N/A Value > Country Median
Lifeclean29 = within(Lifeclean28, 
                     School1 <- 
                       ifelse(is.na(Schooling)=="TRUE",
                              MedianSchoolCountry,
                              Schooling))

#If statement to cascade which median to use: Non-N/A value > Country Median > Continent Median
## Could be more efficient, but couldn't get nested ifelse to work
Lifeclean30 = within(Lifeclean29, 
                     SchoolFinal <- 
                       ifelse(is.na(School1)=="TRUE",
                              MedianSchoolContinent,
                              School1))

#Drop Unneeded Columns
Lifeclean31 = select(Lifeclean30,-c(Schooling,MedianSchoolContinent,MedianSchoolCountry,School1))

##SanityCheck
missmap(Lifeclean31)

##Compute Median by Country
Thin5Country = Lifeclean31 %>% 
  filter(is.na(thinness.5.9.years) == "FALSE") %>% 
  group_by(Country) %>% 
  summarize(MedianThin5Country = median(thinness.5.9.years))

##Compute Median by Continent
Thin5Continent = Lifeclean31 %>% 
  filter(is.na(thinness.5.9.years) == "FALSE") %>% 
  group_by(Continent) %>% 
  summarize(MedianThin5Continent = median(thinness.5.9.years))

#Merge Medians by Continent
Lifeclean32 <- merge(x = Lifeclean31,y = Thin5Continent,by = "Continent",all.x = TRUE)

#Merge Medians by Country
Lifeclean33 <- merge(x = Lifeclean32,y = Thin5Country,by = "Country",all.x = TRUE)

#If statement to cascade which median to use: Non-N/A Value > Country Median
Lifeclean34 = within(Lifeclean33, 
                     Thin51 <- 
                       ifelse(is.na(thinness.5.9.years)=="TRUE",
                              MedianThin5Country,
                              thinness.5.9.years))

#If statement to cascade which median to use: Non-N/A value > Country Median > Continent Median
## Could be more efficient, but couldn't get nested ifelse to work
Lifeclean35 = within(Lifeclean34, 
                     Thin5Final <- 
                       ifelse(is.na(Thin51)=="TRUE",
                              MedianThin5Continent,
                              Thin51))

#Drop Unneeded Columns
Lifeclean36 = select(Lifeclean35,-c(thinness.5.9.years,MedianThin5Continent,MedianThin5Country,Thin51))

##SanityCheck
missmap(Lifeclean36)

##Compute Median by Country
Thin10Country = Lifeclean36 %>% 
  filter(is.na(thinness..1.19.years) == "FALSE") %>% 
  group_by(Country) %>% 
  summarize(MedianThin10Country = median(thinness..1.19.years))

##Compute Median by Continent
Thin10Continent = Lifeclean36 %>% 
  filter(is.na(thinness..1.19.years) == "FALSE") %>% 
  group_by(Continent) %>% 
  summarize(MedianThin10Continent = median(thinness..1.19.years))

#Merge Medians by Continent
Lifeclean37 <- merge(x = Lifeclean36,y = Thin10Continent,by = "Continent",all.x = TRUE)

#Merge Medians by Country
Lifeclean38 <- merge(x = Lifeclean37,y = Thin10Country,by = "Country",all.x = TRUE)

#If statement to cascade which median to use: Non-N/A Value > Country Median
Lifeclean39 = within(Lifeclean38, 
                     Thin101 <- 
                       ifelse(is.na(thinness..1.19.years)=="TRUE",
                              MedianThin10Country,
                              thinness..1.19.years))

#If statement to cascade which median to use: Non-N/A value > Country Median > Continent Median
## Could be more efficient, but couldn't get nested ifelse to work
Lifeclean40 = within(Lifeclean39, 
                     Thin10Final <- 
                       ifelse(is.na(Thin101)=="TRUE",
                              MedianThin10Continent,
                              Thin101))

#Drop Unneeded Columns
Lifeclean41 = select(Lifeclean40,-c(thinness..1.19.years,MedianThin10Continent,MedianThin10Country,Thin101))

##SanityCheck
missmap(Lifeclean41)

##Compute Median by Country
Thin10Country = Lifeclean36 %>% 
  filter(is.na(thinness..1.19.years) == "FALSE") %>% 
  group_by(Country) %>% 
  summarize(MedianThin10Country = median(thinness..1.19.years))

##Compute Median by Continent
Thin10Continent = Lifeclean36 %>% 
  filter(is.na(thinness..1.19.years) == "FALSE") %>% 
  group_by(Continent) %>% 
  summarize(MedianThin10Continent = median(thinness..1.19.years))

#Merge Medians by Continent
Lifeclean37 <- merge(x = Lifeclean36,y = Thin10Continent,by = "Continent",all.x = TRUE)

#Merge Medians by Country
Lifeclean38 <- merge(x = Lifeclean37,y = Thin10Country,by = "Country",all.x = TRUE)

#If statement to cascade which median to use: Non-N/A Value > Country Median
Lifeclean39 = within(Lifeclean38, 
                     Thin101 <- 
                       ifelse(is.na(thinness..1.19.years)=="TRUE",
                              MedianThin10Country,
                              thinness..1.19.years))

#If statement to cascade which median to use: Non-N/A value > Country Median > Continent Median
## Could be more efficient, but couldn't get nested ifelse to work
Lifeclean40 = within(Lifeclean39, 
                     Thin10Final <- 
                       ifelse(is.na(Thin101)=="TRUE",
                              MedianThin10Continent,
                              Thin101))

#Drop Unneeded Columns
Lifeclean41 = select(Lifeclean40,-c(thinness..1.19.years,MedianThin10Continent,MedianThin10Country,Thin101))

##SanityCheck
missmap(Lifeclean41)

##Compute Median by Country
BMICountry = Lifeclean41 %>% 
  filter(is.na(BMI) == "FALSE") %>% 
  group_by(Country) %>% 
  summarize(MedianBMICountry = median(BMI))

##Compute Median by Continent
BMIContinent = Lifeclean41 %>% 
  filter(is.na(BMI) == "FALSE") %>% 
  group_by(Continent) %>% 
  summarize(MedianBMIContinent = median(BMI))

#Merge Medians by Continent
Lifeclean42 <- merge(x = Lifeclean41,y = BMIContinent,by = "Continent",all.x = TRUE)

#Merge Medians by Country
Lifeclean43 <- merge(x = Lifeclean42,y = BMICountry,by = "Country",all.x = TRUE)

#If statement to cascade which median to use: Non-N/A Value > Country Median
Lifeclean44 = within(Lifeclean43, 
                     BMI1 <- 
                       ifelse(is.na(BMI)=="TRUE",
                              MedianBMICountry,
                              BMI))

#If statement to cascade which median to use: Non-N/A value > Country Median > Continent Median
## Could be more efficient, but couldn't get nested ifelse to work
Lifeclean45 = within(Lifeclean44, 
                     BMIFinal <- 
                       ifelse(is.na(BMI1)=="TRUE",
                              MedianBMIContinent,
                              BMI1))

#Drop Unneeded Columns
Lifeclean46 = select(Lifeclean45,-c(BMI,MedianBMIContinent,MedianBMICountry,BMI1))

##SanityCheck
missmap(Lifeclean46)

##Compute Median by Country
DiptheriaCountry = Lifeclean46 %>% 
  filter(is.na(Diphtheria) == "FALSE") %>% 
  group_by(Country) %>% 
  summarize(MedianDiptheriaCountry = median(Diphtheria))

##Compute Median by Continent
DiptheriaContinent = Lifeclean46 %>% 
  filter(is.na(Diphtheria) == "FALSE") %>% 
  group_by(Continent) %>% 
  summarize(MedianDiptheriaContinent = median(Diphtheria))

#Merge Medians by Continent
Lifeclean47 <- merge(x = Lifeclean46,y = DiptheriaContinent,by = "Continent",all.x = TRUE)

#Merge Medians by Country
Lifeclean48 <- merge(x = Lifeclean47,y = DiptheriaCountry,by = "Country",all.x = TRUE)

#If statement to cascade which median to use: Non-N/A Value > Country Median
Lifeclean49 = within(Lifeclean48, 
                     Dip1 <- 
                       ifelse(is.na(Diphtheria)=="TRUE",
                              MedianDiptheriaCountry,
                              Diphtheria))

#If statement to cascade which median to use: Non-N/A value > Country Median > Continent Median
## Could be more efficient, but couldn't get nested ifelse to work
Lifeclean50 = within(Lifeclean49, 
                     DiptheriaFinal <- 
                       ifelse(is.na(Dip1)=="TRUE",
                              MedianDiptheriaContinent,
                              Dip1))

#Drop Unneeded Columns
Lifeclean51 = select(Lifeclean50,-c(Diphtheria,MedianDiptheriaContinent,MedianDiptheriaCountry,Dip1))

##SanityCheck
missmap(Lifeclean51)

##Compute Median by Country
PolioCountry = Lifeclean51 %>% 
  filter(is.na(Polio) == "FALSE") %>% 
  group_by(Country) %>% 
  summarize(MedianPolioCountry = median(Polio))

##Compute Median by Continent
PolioContinent = Lifeclean51 %>% 
  filter(is.na(Polio) == "FALSE") %>% 
  group_by(Continent) %>% 
  summarize(MedianPolioContinent = median(Polio))

#Merge Medians by Continent
Lifeclean52 <- merge(x = Lifeclean51,y = PolioContinent,by = "Continent",all.x = TRUE)

#Merge Medians by Country
Lifeclean53 <- merge(x = Lifeclean52,y = PolioCountry,by = "Country",all.x = TRUE)

#If statement to cascade which median to use: Non-N/A Value > Country Median
Lifeclean54 = within(Lifeclean53, 
                     Polio1 <- 
                       ifelse(is.na(Polio)=="TRUE",
                              MedianPolioCountry,
                              Polio))

#If statement to cascade which median to use: Non-N/A value > Country Median > Continent Median
## Could be more efficient, but couldn't get nested ifelse to work
Lifeclean55 = within(Lifeclean54, 
                     PolioFinal <- 
                       ifelse(is.na(Polio1)=="TRUE",
                              MedianPolioContinent,
                              Polio1))

#Drop Unneeded Columns
Lifeclean56 = select(Lifeclean55,-c(Polio,MedianPolioContinent,MedianPolioCountry,Polio1))

##SanityCheck
missmap(Lifeclean56)

#Change Column Names back to Original
names(Lifeclean56) <- c("Country","Continent","Year","Status","Life.expectancy","Adult.Mortality","infant.deaths",
                        "percentage.expenditure","Measles","under.five.deaths","HIV.AIDS","Population","GDP",
                        "Hepatitis.B","Total.expenditure","Alcohol","Income.composition.of.resources",
                        "Schooling","thinness.5.9.years","thinness..1.19.years","BMI","Diphtheria","Polio")

#Create Final Dataset
LifeFinal = Lifeclean56

#Final Sanity Check - only N/A's relate to South Sudan's GDP (Wasn't a country prior to 2008). All set!
missmap(LifeFinal)
LifeFinal %>% filter(is.na(GDP)=="TRUE")
```


##Life Expectancy Random Forest

```{r Create the Random Forest Model}
#Load Libraries
library(randomForest)
library(tree)

#Remove N/A's for modeling
LifeFinal1 = LifeFinal %>% filter(is.na(GDP) == "FALSE")

#Run Model
set.seed(1)
train = sample(1:nrow(LifeFinal1),nrow(LifeFinal1)/2)
life.test = LifeFinal1[-train,"Life.expectancy"]
rf.life = randomForest(Life.expectancy~.,data = LifeFinal1,subset=train,mtry=5,importance=TRUE)
yhat.rf=predict(rf.life,newdata = LifeFinal1[-train,])
sqrt(mean((yhat.rf-life.test)^2))

#see importance of variables
importance(rf.life)
varImpPlot(rf.life,main = 'Variable Importance for Life Expectancy')

#See the Full Model
rf.life
```

